Dark Users
========================================================

Quantify the extent to which user identifiers left in activity traces (forums so far) do not appear in the hash_mapping and users tables. Main upshot is that we lack user attributes, although we could probably assume the dark users are withdrawn students and so impute access_group_id=4 without much error.

**gen.users and **map.hash_mapping are believed to correspond 1:1

The following considers forum_user_ids in threads, posts and comments. "Including dark" counts forum users as identified in the forum-specific tables, whereas "excluding dark" only counts those users with their forum_user_id also found in the hash_mapping table.

```{r echo=FALSE}
courseIDs<-c("aiplan","astro","crit","edc","equine","intro")
echo.sql<-TRUE # echo SQL statements
#establish connection to MySQL, loading library. contains coursera DB exports from 2013
source("../dbConnect.R")
#helper functions
source("../helpers.R")

includingDark.sql<-"SELECT count(1) including_dark FROM (
    (SELECT DISTINCT fp.forum_user_id from **for.forum_posts fp)
    UNION DISTINCT
    (SELECT DISTINCT fc.forum_user_id from **for.forum_comments fc)
    UNION DISTINCT
    (SELECT DISTINCT ft.forum_user_id from **for.forum_threads ft) ) i"
excludingDark.sql<-"SELECT count(1) excluding_dark FROM (
    (SELECT DISTINCT fp.forum_user_id from **for.forum_posts fp, **map.hash_mapping m
        WHERE fp.forum_user_id = m.forum_user_id)
    UNION DISTINCT
    (SELECT DISTINCT fc.forum_user_id from **for.forum_comments fc, **map.hash_mapping m
        WHERE fc.forum_user_id = m.forum_user_id)
    UNION DISTINCT
    (SELECT DISTINCT ft.forum_user_id from **for.forum_threads ft, **map.hash_mapping m
        WHERE ft.forum_user_id = m.forum_user_id) ) i"
forumsDark.df<-tabular.SELECT(db, courseIDs, includingDark.sql, echo=echo.sql)
forumsDark.df<-cbind(forumsDark.df,tabular.SELECT(db, courseIDs, excludingDark.sql, echo=echo.sql))
pc.df<-data.frame(percent_lost=100-round(100*forumsDark.df[,"excluding_dark"]/forumsDark.df[,"including_dark"]))
forumsDark.df<-cbind(forumsDark.df,pc.df)
```

```{r echo=FALSE, results='asis'}
kable(forumsDark.df, row.names=T, align=rep('c',3))
```

