```{r echo=FALSE, results='hide', message=FALSE}
if(!exists("courseID", inherits=T)){
   #specify a single course here. Ignored when called (iteratively) from "Degree P-C - All Courses.R"
    courseID<-"equine"#c("aiplan","astro","crit","edc","equine","intro")
}
top.quantile<-0.0025 #0.005 means look at posters in the top 0.5% of degree distribution
net.type="P-C" #poster-commenter filename prefix
store.dir<-"~/R Projects/Edinburgh MOOC/EdMOOC-SNA/Network Data/" #where the extracted networks are stored
echo.flag<-TRUE #echo R source
fig.width.default<-5
fig.height.default<-5
options(knitr.table.format = 'markdown')
library("sna")
library("network")
```

Extract and Visualise the Sub-network Containing the Highest In- and Out-degree Individuals - `r courseID`
========================================================

This treats "Poster-Commenter" networks. The assumption of the following that a tie between individuals is defined by a comment on a post. This is a directed tie from the commenter to the poster. A tie is assumed if there are 1 or more comments and ties are **not** weighted by the number of comments.


```{r  echo=FALSE}
#loads RData file containiing network object and metadata
load(paste(store.dir,net.type," ",courseID,".RData", sep=""))
cat(paste("Loaded data. Created by:",metadata$origin,"Created on:",metadata$created))
print(net1)
```

## Preliminaries - identify the highest in/out degree individuals

```{r  echo=echo.flag}
odeg<-degree(net1, cmode="outdegree")
ideg<-degree(net1, cmode="indegree")
#get indexes to nodes and filter for tiny quantile (top.quantile)
# not happy with this method, what is the best cut off to give comparable views between courses?
ocentile<-quantile(odeg,1-top.quantile)
icentile<-quantile(ideg,1-top.quantile)
oindex<-which(odeg>=ocentile)
iindex<-which(ideg>=icentile)
odeg<-odeg[oindex]
ideg<-ideg[iindex]
#order
oorder<-order(odeg, decreasing=T)
iorder<-order(ideg, decreasing=T)
odeg<-odeg[oorder]
ideg<-ideg[iorder]
oindex<-oindex[oorder]
iindex<-iindex[iorder]
# combine forum_user_id and role from nodes with degree.
vroles<-net1 %v% "role"
vnames<-net1 %v% "vertex.names"
outdegree.df<-cbind(forum_user_id=vnames[oindex], role=vroles[oindex], out.degree=odeg, match.in=oindex %in% iindex)
indegree.df<-cbind(forum_user_id=vnames[iindex], role=vroles[iindex], in.degree=ideg, match.out=iindex %in% oindex)
```

**People with highest `r top.quantile*100`% Quantile Out-degree**

```{r results='asis', echo=FALSE}
kable(outdegree.df)
```

**People with Highest `r top.quantile*100`% Quantile In-degree**

```{r results='asis', echo=FALSE}
kable(indegree.df)
```
_Match.in/out shows where the same person appears in the other list._

## Extract the Subnet

This includes all of the people above and any other person who they commented on or who commented on them. Individuals with 1 or only a few ties are not from the above sets but have been included as due to a tie with at least 1 member of these lists.
```{r fig.width=10, fig.height=10}
# obtain the subgraph containing these nodes and linked edges.
# NB: get.neighbourhood is not vectorised!
o.connected<-c(oindex,unlist(sapply(oindex, function(x){get.neighborhood(net1,x, type="combined")})))
i.connected<-c(iindex,unlist(sapply(iindex, function(x){get.neighborhood(net1,x, type="combined")})))
connected<-unique(c(o.connected,i.connected))
top.subnet<-net1 %s% connected
vertex.col<-as.numeric(top.subnet %v% "role")
# we could colourise according to whether people were in the top set or only connected to them
# using vertex.col<-as.numeric(connected[order(connected)] %in% c(oindex,iindex))+1 #order needed!
plot(top.subnet, arrowhead.cex=0.7, edge.lwd=0.4, vertex.col=vertex.col, vertices.last=FALSE)
```
There are `r length(connected)` individuals in the diagram.

In practice, choosing selecting more high degree people selected, say 0.5% vs 0.25%, does not make much difference to the appearance of the subnet if there are 1 or 2 very high degree individuals.

### The Network of Associates

What does the network look like if we take away the highest degree people and only look at the people they were connected to (who were not already in either list, above).

```{r fig.width=10, fig.height=10}
connected.associates<-connected[!connected %in% c(oindex,iindex)]
associates.subnet<-net1 %s% connected.associates
vertex.col<-as.numeric(associates.subnet %v% "role")
# we could colourise according to whether people were in the top set or only connected to them
# using vertex.col<-as.numeric(connected[order(connected)] %in% c(oindex,iindex))+1 #order needed!
plot(associates.subnet, arrowhead.cex=0.7, edge.lwd=0.4, vertex.col=vertex.col, vertices.last=FALSE)
```
There are `r length(connected.associates)` individuals in the diagram.


### Compare to a randomly selected group of people numbering the same as the "associates"

_The number of people is only approximately the same because `r length(connected.associates)` are drawn with replacement and then unique people plotted. Should be good enough for a qualitative comparison, which is what is intended, and the fraction of the people available is small._

Beware: for a better comparison, we should randomly select people (posters) who have been commented on, or who have commented.

```{r fig.width=10, fig.height=10}
connected.random<-unique(round(runif(length(connected.associates),1,network.size(net1))))
random.subnet<-net1 %s% connected.random
vertex.col<-as.numeric(random.subnet %v% "role")
plot(random.subnet, arrowhead.cex=0.7, edge.lwd=0.4, vertex.col=vertex.col, vertices.last=FALSE)
```


## Extract the subnet for only the highest degree (in and out) person and their ties.
This will select 1 or 2 people and their ties, depending on whether the same individual has the highest in and out degrees or not.
```{r fig.width=10, fig.height=10}
o.connected.1<-c(oindex[1],get.neighborhood(net1,oindex[1], type="combined"))
i.connected.1<-c(iindex[1],get.neighborhood(net1,iindex[1], type="combined"))
connected.1<-unique(c(o.connected.1,i.connected.1))
top.1.subnet<-net1 %s% connected.1
vertex.1.col<-as.numeric(top.1.subnet %v% "role")
plot(top.1.subnet, arrowhead.cex=0.7, edge.lwd=0.4, vertex.col=vertex.1.col, vertices.last=FALSE)
```
There are `r length(connected.1)` individuals in the diagram.

```{r fig.width=4, fig.height=5, echo=F}
plot.new()
leg=c("Administrator","Instructor","Teaching Staff","Student","Blocked","Student Access","Community TA","School Administrator","Student (Forum Banned)")
legend(x=0,y=1,title="Roles", legend=leg, seq(1,9))
```

```{r echo=FALSE}
#save in UCINET DL format for gephi. This is limited to a single vertex label AFAIK.
# There are self-loops in exported data, so I assume loops=FALSE network attribute controls processing, not the data structure
fname<-paste("~/R Projects/Edinburgh MOOC/EdMOOC-SNA/Top Subnets/Degree ", net.type, " - " ,courseID,".dl",sep="")
write.dl(top.subnet, fname, vertex.lab=top.subnet %v% "role")
cat(paste("Saved UCINET DL file to:", fname))
```


```{r echo=FALSE}
## ****************************************************
## Created: Adam Cooper, Cetis, Oct 2013
## This source code was produced for The University of
## Edinburgh DEI as part of their MOOC initiative.
## ****************************************************

## ***Made available using the The MIT License (MIT)***
#The MIT License (MIT)
#Copyright (c) 2013 Adam Cooper, University of Bolton
#
#Permission is hereby granted, free of charge, to any person obtaining a copy of
#this software and associated documentation files (the "Software"), to deal in
# the Software without restriction, including without limitation the rights to
# use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of
# the Software, and to permit persons to whom the Software is furnished to do so,
# subject to the following conditions:
#    
#    The above copyright notice and this permission notice shall be included in all
# copies or substantial portions of the Software.
# 
# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS
# FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR
# COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER
# IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN
# CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
## ************ end licence ***************
```