Basic Forum Stats
========================================================

```{r echo=FALSE, results='hide'}
courseIDs<-c("aiplan","astro","crit","edc","equine","intro")
echo.sql<-TRUE # echo SQL statements
echo.flag<-TRUE #echo R source
fig.width.default<-5
fig.height.default<-5
options(knitr.table.format = 'markdown')
read_chunk("./Basic Forum Stats.R")
```

```{r INIT, echo=FALSE}
#initialises the DB and sources helper.
```

Quantify users vs registrations with "Student" role (access_group_id=4).
```{r ACCESSED, echo=FALSE}
```
```{r echo=FALSE, results='asis'}
kable(accessed.df, row.names=T, align='c')
```

Proportion of show/no-show.
```{r echo=echo.flag, fig.width=fig.width.default, fig.height=fig.height.default}
accessed.proportions<-t(accessed.df$accessed/accessed.df$registered)
colnames(accessed.proportions)<-row.names(accessed.df)
barplot(rbind(accessed=accessed.proportions, no.show=1-accessed.proportions), main="Registrations with at Least 1 Access Event", ylab="Conversion Fraction")

```


Qty of forums, threads, posts, comments
-----------

```{r FORUM_COUNT, echo=FALSE}
```
### Forums

```{r echo=FALSE, results='asis'}
#curious align due to bug in knitr
kable(forumCount.df, row.names=T, align=rep("c", length(colnames(forumCount.df))))
```

### Threads

```{r echo=FALSE, results='asis'}
kable(threadCount.df, row.names=T, align=rep("c", length(colnames(threadCount.df))))
```

This does not count threads in a deleted parent forum as being deleted threads.

### Posts

```{r echo=FALSE, results='asis'}
kable(postCount.df, row.names=T, align=rep("c", length(colnames(postCount.df))))
```

This does not count posts in a deleted parent thread as being deleted posts.

### Comments

```{r echo=FALSE, results='asis'}
kable(commentCount.df, row.names=T, align=rep("c", length(colnames(commentCount.df))))
```

This does not count comments in a deleted parent posts as being deleted comments.

### Orphan Posts and Comments

Orphan posts are posts with thread or forum parents that are deleted (and the post is not also deleted).
Orphan comments are similarly defined, with "orphan comments (parent)" counting comments whose parent post is deleted whereas "orphan comments" are those missing an ancestor going back up to the forum level.

```{r FORUM_ORPHANS, echo=FALSE}
```
```{r echo=FALSE, results='asis'}
kable(orphans.df, row.names=T, align=rep("c", length(colnames(commentCount.df))))
```

What are Forums For? Also Thread-count and and Timing
-------------
```{r FORUM_TOP_STRUCT, echo=FALSE}
```
```{r echo=FALSE, results='asis'}

for(i in 1:length(forumTopStruct)){
   cat(paste("### ",courseIDs[i],"\r\n\r\n"))
   kable(forumTopStruct[[i]], row.names=F, align=rep("c", length(colnames(commentCount.df))))
}

```

Deleted forums are excluded. In practice, these are all threadless, except 1 case in aiplan with 1 thread.

No of posters by role
-------------

```{r POSTER_ROLES, echo=FALSE}
```
```{r echo=FALSE, results='asis',}
kable(posterRoles.df, row.names=T, align=rep("c", length(colnames(posterRoles.df))))
```
*role names are assumed to be the same for the same id across courses*

Deleted posts and posts in deleted threads or forums are not counted. Not counting posters to deleted forums or threads reduces the count by up to around 1%.

For the student role, what fraction of the registered students who logged in at least once does this represent?
```{r echo=echo.flag}
#data frame rows will be in same order as schemaIDs
stuPostFrac<-t(posterRoles.df[,"Student"]/accessed.df[,"accessed"])
colnames(stuPostFrac)<-row.names(posterRoles.df)
```
```{r echo=FALSE, results='asis',}
kable(stuPostFrac, row.names=T)
```

Distribution of No. of Posts Per Person
----------
What is the distribution of the number of people making 1,2,... posts? Deleted posts and posts to deleted threads/forums are excluded. This is probably of little consequence, but there may be cases of tutors posting tests or other aftermath posts with no relevance to the student's experience.

*Non-posters are excluded*
```{r POSTER_DISTRIBUTION, echo=FALSE}
```

### Students (access_group_id=4)

```{r echo=echo.flag, fig.width=fig.width.default, fig.height=fig.height.default}
# remove cases where no of posts >20 to simplify plotting
for(i in 1:length(posterDistS)){
   posterDistSa<-posterDistS[[i]][posterDistS[[i]]$posts<=20,"posts"]
   cut.posters<-length(posterDistS[[i]]$posts[posterDistS[[i]]$posts>20])
   hist(posterDistSa, main=names(posterDistS)[i], xlab="No. of Posts", ylab="No. of People", sub=paste(cut.posters,"posters with >20 posts"), breaks=20)
}
```

I suspect that there are quite a few 1-post people due to a different mechanism - e.g. "introduce yourself" - compared to people who actually used the forums for a purpose. Hence if we wanted to compare distributions between courses we should probably cut 1 (or even 1-2) posters. These should appear (mostly) in 1 forum of thread, so could be eliminated if forum/thread titles inspected. Alt, if mixed in, could estimate % by sampling and manual classification.

### Teaching Staff (access_group_id IN (2,3))

```{r echo=echo.flag, fig.width=fig.width.default, fig.height=fig.height.default}

for(i in 1:length(posterDistT)){
   hist(posterDistT[[i]]$posts, main=names(posterDistT)[i], xlab="No. of Posts", ylab="No. of People")
}
```

Distribution of No. of Comments Per Person
----------
What is the distribution of the number of people making 1,2,... comments? Comments on deleted posts *are* counted this time since they probably do represent something meaningful about the students' experience of the course.

*Non-commenters are excluded*
```{r COMMENTER_DISTRIBUTION, echo=FALSE}
```

### Students (access_group_id=4)

```{r echo=echo.flag, fig.width=fig.width.default, fig.height=fig.height.default}
# remove cases where no of posts >20 to simplify plotting
for(i in 1:length(posterDistS)){
   commenterDistSa<-commenterDistS[[i]][commenterDistS[[i]]$comments<=15,"comments"]
   cut.commenters<-length(commenterDistS[[i]]$comments[commenterDistS[[i]]$comments>15])
   hist(commenterDistSa, main=names(commenterDistS)[i], xlab="No. of Comments", ylab="No. of People", sub=paste(cut.commenters,"people with >15 comments"), breaks=15)
}
```

### Teaching Staff (access_group_id IN (2,3))

```{r echo=echo.flag, fig.width=fig.width.default, fig.height=fig.height.default}
for(i in 1:length(posterDistT)){
   hist(commenterDistT[[i]]$comments, main=names(commenterDistT)[i], xlab="No. of Comments", ylab="No. of People")
}
```

Density comments on posts (etc), compare distributions
-----------

### Bushiness Factors

```{r FORUM_BUSHINESS, echo=FALSE}
```
Compute the degree - i.e. the average number of "child" entities per parent - to which forums, threads, post and comments branch. The root forum (with parent=-1) is not counted. Deleted entities and entities with deleted ancestors are not counted. Note that some course have forums containing forums, and threads in both levels of forum. This relationship between forums is ignored; each forum is counted as "1".

Excluding deletions makes <10% difference to post and comment bushiness but changes thread bushiness by several tens of % since some course data has a lot of forum deletion.

```{r echo=FALSE, results='asis'}
kable(bushiness.df, row.names=T, align=rep("c", length(colnames(commentCount.df))))
```

### Distribution of Comments on Posts

Comments on deleted posts are not included. This does make a visible difference to the following bar plot some courses, notably "aiplan", for which the proportion of no-comment posts drops by around 3%. 

```{r FORUM_DISTRIBUTION, echo=FALSE}
```

```{r echo=echo.flag, fig.width=fig.width.default, fig.height=fig.height.default}
barplot(100*post_0comments/postCount.df$posts, main="Proportion of posts with no comments", ylab="%")
```

What is the distribution of the number of posts with 1,2,... comments on them?

```{r echo=echo.flag, fig.width=fig.width.default, fig.height=fig.height.default}
# remove cases where comments>20 to simplify plotting
for(i in 1:length(post_commentDist2)){
   post_commentDist2a<-post_commentDist2[[i]][post_commentDist2[[i]]$comments<=20,"comments"]
   cut.posts<-length(post_commentDist2[[i]]$comments[post_commentDist2[[i]]$comments>20])
   hist(post_commentDist2a, main=names(post_commentDist2)[i], xlab="No. of Comments", ylab="No. of Posts", sub=paste(cut.posts,"posts with >20 comments"))
}
```


```{r EXIT, echo=FALSE}
#saves extracted data to an RData file and closes DB connection
```
```{r echo=FALSE}
## ****************************************************
## Created: Adam Cooper, Cetis, Oct 2013
## This source code was produced for The University of
## Edinburgh DEI as part of their MOOC initiative.
## ****************************************************

## ***Made available using the The MIT License (MIT)***
#The MIT License (MIT)
#Copyright (c) 2013 Adam Cooper, University of Bolton
#
#Permission is hereby granted, free of charge, to any person obtaining a copy of
#this software and associated documentation files (the "Software"), to deal in
# the Software without restriction, including without limitation the rights to
# use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of
# the Software, and to permit persons to whom the Software is furnished to do so,
# subject to the following conditions:
#    
#    The above copyright notice and this permission notice shall be included in all
# copies or substantial portions of the Software.
# 
# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS
# FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR
# COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER
# IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN
# CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
## ************ end licence ***************
```