Forum Activity By Region
========================================================

**TO DO: check consistency of calculations, smarten chart labels and legends, add comment calculations? (maybe check significance of difference to posting - if none, consider merging posts and comments as "contibutions"..**

```{r echo=FALSE, results='hide'}
courseIDs<-c("aiplan","astro","crit","edc","equine","intro")
echo.sql<-TRUE # echo SQL statements
echo.flag<-TRUE #echo R source
fig.width.default<-7
fig.height.default<-7
options(knitr.table.format = 'markdown')
read_chunk("./Regional Differences.R")
```

```{r INIT, echo=FALSE, message=FALSE}
#initialises the DB and sources helper.
```

Investigate how people from different regions participate in the forums. Only people in the "student" role (access_group_id=4) are considered. Withdrawn/deleted users are neglected (role and region are unknown).

Regions are derived from the continent and country determined from IP address and comprise a crude proxy for culture and native language. Statistics on English-speaking are not used, but see: http://en.wikipedia.org/wiki/List_of_countries_by_English-speaking_population. The default region is a continent, but the following exceptions are made to this rule:
* UK AND IRELAND designates tha UK, Ireland and the Isle of Man
* EUROPE is the rest of Europe, including countries outside the EU
* CHINA includes Hong Kong and Taiwan
* INDIA is separated from other Asian countries
* AUSTRALIA AND NZ are separated off from Oceania 
* Mexico is recoded to CENTRAL AMERICA
* The Caribbean is recoded to CENTRAL AMERICA (small numbers)

The "introductions" forums are considered separately since they have quite a different character to other forums.

Regional Distribution of Users
----------

Although we are mostly concerned with differences in behaviour, it is also interesting to get a feel for whether or not this is likely to influence the overall forum character. Hence the proportion of users from the regions is of interest.
```{r USER_DIST, echo=FALSE}
```
```{r echo=FALSE, results='asis'}
kable(userRegions.df, row.names=T, align='c')
```

```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
cols<-rainbow(length(userRegions.df[1,]))
par(mar=c(5.1, 4.1, 4.1, 7.1), xpd=TRUE)
barplot(100*t(as.matrix(userRegions.df)/rowSums(userRegions.df)), ylab="% of participants", col=cols, las=2)
legend("topright", inset=c(-0.25,0),legend=rev(colnames(userRegions.df)), fill=rev(cols), cex=0.6)
```

Probability of Contributing By Region
------

Since the number of users differs widely between regions, we will look at the probability that a given user will contribute either a comment or a post. 

```{r echo=T}
# specify which forum ids are for personal introductions
intro.fid<-c(aiplan=10,astro=12,crit=2,edc=25,equine=2,intro=11)
# and for technical queries
tech.fid<-c(aiplan=9,astro=9,crit=9,edc=20,equine=9,intro=9)
# all other forums are assumed to be subject matter
```

## Probability of Posting

What is the probability that a randomly chosen person from a given region in each course will have contributed at least 1 post.

```{r POST_PROBS, echo=FALSE}
```

A "heatmap" visualises the hot-spots in this table, where paler colours indicate which region has people who are more likely to post. This is relative to other regions; colouring should be read across rows. The branching diagrams (dendrograms) indicate relative similarity based on these values as judged by a process of hierarchical clustering. Similarity is judged taking all courses into account when clustering by region, and by taking all regions into account (equally-weighted) when clustering by course. For example, this indicates that India and China similar and distinct from the rest of the world. The regional dendrogram also appears to show similarity among regions for which English is predominantly the first language.

```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
heatmap(as.matrix(postProb.subj), main="Subject Forum Posting Probability")
```

Look at each course separately because the differing levels of forum use between courses, and the quantity of data-points, confuses matters if courses are plotted side-by-side on the same bar plot.

### AI Planning
```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
plot.1course("aiplan")
```

**The numbers of people in some regions is quite small; be cautious about interpretation.**

### Astro-biology
```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
plot.1course("astro")
```

### Critical Thinking
```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
plot.1course("crit")
```

### eL & DC
```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
plot.1course("edc")
```

### Equine Nutririon
```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
plot.1course("equine")
```

### Intro to Philo
```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
plot.1course("intro")
```


### Significance

Given the sample sizes, are these differences statistically significant? Consider only the subject-matter forum posts
```{r BINOM_TESTS, echo=FALSE, message=FALSE}
```
The following table shows the percentage chance that the number of posters can be assumed to be the same as the average for the course. A value less than 5% would be a conventional threshold for asserting that the number of people posting in that region is significantly different from what would be exected from the whole-course average. Whether it is higher or lower than the mean can be judged from previous charts.

```{r echo=FALSE, results='asis'}
kable(100*H0.probs, row.names=T, align='c')
```


## Average Number of Posts per Person

Average posts per person (AI planning excluded) in the subject-matter forums.
```{r POSTS_PER_PERSON, echo=FALSE}
```
```{r echo=FALSE, results='asis'}
postsPerPerson.subj.x<-postsPerPerson.subj[-1,]
kable(postsPerPerson.subj.x, row.names=T, align='c')
```



```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
heatmap(as.matrix(postsPerPerson.subj.x))
```

### Comparing Courses

```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
cols<-rainbow(length(postsPerPerson.subj.x[1,]))
par(mar=c(5.1, 4.1, 4.1, 7.1), xpd=TRUE)
barplot(t(as.matrix(postsPerPerson.subj.x)), ylab="Average Posts per Person", col=cols, las=2, beside=T)
legend("topright", inset=c(-0.15,0),legend=rev(colnames(userRegions.df)), fill=rev(cols), cex=0.6)
```

### Comparing Regions

```{r echo=echo.flag, fig.height=fig.height.default, fig.width=fig.width.default}
cols<-rainbow(length(postsPerPerson.subj.x[,1]))
par(mar=c(5.1, 4.1, 4.1, 7.1), xpd=TRUE)
barplot(as.matrix(postsPerPerson.subj.x), ylab="Average Posts per Person", col=cols, las=2, beside=T)
legend("topright", inset=c(-0.15,0),legend=rev(rownames(userRegions.df)), fill=rev(cols), cex=0.6)
```


Regional Distribution of Most Active Users
---------

**could also compare the regional split in the top decile of posters compared to the overall regional split**

Other Ideas
----------
Could repeat for thread-reads. I expect little variation in reads:people between regions. Differences in post:read ratio are effectively already shown above.

```{r EXIT, echo=FALSE, message=FALSE}
#saves extracted data to an RData file and closes DB connection
```
```{r echo=FALSE}
## ****************************************************
## Created: Adam Cooper, Cetis, Nov 2013
## This source code was produced for The University of
## Edinburgh DEI as part of their MOOC initiative.
## ****************************************************

## ***Made available using the The MIT License (MIT)***
#The MIT License (MIT)
#Copyright (c) 2013 Adam Cooper, University of Bolton
#
#Permission is hereby granted, free of charge, to any person obtaining a copy of
#this software and associated documentation files (the "Software"), to deal in
# the Software without restriction, including without limitation the rights to
# use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of
# the Software, and to permit persons to whom the Software is furnished to do so,
# subject to the following conditions:
#    
#    The above copyright notice and this permission notice shall be included in all
# copies or substantial portions of the Software.
# 
# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS
# FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR
# COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER
# IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN
# CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
## ************ end licence ***************
```