```{r echo=FALSE, results='hide'}
#this is normally called from "Tie Type Correlation - All Courses.R"
if(!exists("courseID")){
   courseID<-"astro"#"aiplan","astro","crit","edc","equine","intro"
   tie.type.1<-"P-C"
   tie.type.2<-"P-Set"
}

echo.flag<-TRUE #echo R source
fig.width.default<-5
fig.height.default<-5
options(knitr.table.format = 'markdown')
store.dir<-"~/R Projects/Edinburgh MOOC/EdMOOC-SNA/Network Data/" #where to load data from

library("igraph")
library("lattice")
library("iplots")

ig1<-read.graph(paste(store.dir,tie.type.1," ", courseID,".graphml",sep=""), format="graphml")
ig2<-read.graph(paste(store.dir,tie.type.2," ", courseID,".graphml",sep=""), format="graphml")
```

Analysis of Correlation Between Centrality Measures in Networks with Different Tie Definitions - `r courseID`
========================================================

Instances are individuals.  The two tie types are poster-commenter and post-set.

The attributes are (all undirected measures since P-Set is undirected):
* Degree centrality
* Betweenness centrality
* Eigenvector centrality

Closeness centrality is not used because the values show little spread (in practice, only a few distinct values are seen).

Data for the whole course run is used. This means that we should expect people who dropped out but who did not withdraw to have low centrality measures. People who withdrew do not appear.

Square root transformation is applied to the centrality measures to compress their long-tails.

Isolates are removed. Specifically, isolates in "post-set" are removed. These are people who post and who got no comments, and never comment on anyone else's posts. There may be quite a lot of these and including them distorts attempts to assess correlation. 

There are a small number of people who have zero degree in P-C but non-zero in P-Set (1-2% by inspection). This is thought to arise from cases where a poster withdrew and was removed from the users table. Edges to deleted people are removed (we cannot know their role or geographical region) but the effect of this differs for different tie type definitiions. For P-C, it can lead to commenters who were connected to a withdrawn poster being left as isolates. The same will happen for P-Set when there is only a single commenter, but when there is more than one non-withdrawn commenter, these commenters WILL have a tie. Hence they will have non-zero degree in exactly the same posting-commenting scenario.

```{r echo=echo.flag, message=FALSE}
summary(ig1)
summary(ig2)
```

For comparison, a simulated network is generated with the same number of nodes and edges using Erdos-Renyi generation rules. The generator rules are that edges are randomly distributed.
```{r echo=echo.flag, message=FALSE}
ig3<-erdos.renyi.game(vcount(ig1), ecount(ig1), type="gnm", directed=T)
ig4<-erdos.renyi.game(vcount(ig2), ecount(ig2), type="gnm", directed=F)
```

```{r echo=echo.flag}
sim1.df<-data.frame(sqrt_degree=sqrt(degree(ig3, mode="all", loops=FALSE)),
                 sqrt_betweenness=sqrt(betweenness(ig3, directed=FALSE)),
                  sqrt_eigenvector=sqrt(evcent(ig3, directed=FALSE)$vector))
plot(sim1.df, main="Correlation of Square Root of Centrality Measures in Random Graph", sub="Poster-Commenter")
sim2.df<-data.frame(sqrt_degree=sqrt(degree(ig4, mode="all", loops=FALSE)),
                 sqrt_betweenness=sqrt(betweenness(ig4, directed=FALSE)),
                  sqrt_eigenvector=sqrt(evcent(ig4, directed=FALSE)$vector))
plot(sim2.df, main="Correlation of Square Root of Centrality Measures in Random Graph", sub="Post-Set")
```

```{r echo=echo.flag}
#build a dataframe for person-level centrality measures+role (for labelling)
m1.df<-data.frame(role=get.vertex.attribute(ig1, name="role"),
                 sqrt_degree1=sqrt(degree(ig1, mode="all", loops=FALSE)),
                 sqrt_betweenness1=sqrt(betweenness(ig1, directed=FALSE)),
                  sqrt_eigenvector1=sqrt(evcent(ig1, directed=FALSE)$vector))
m2.df<-data.frame(sqrt_degree2=sqrt(degree(ig2, mode="all", loops=FALSE)),
                 sqrt_betweenness2=sqrt(betweenness(ig2, directed=FALSE)),
                  sqrt_eigenvector2=sqrt(evcent(ig2, directed=FALSE)$vector))
#ensure that the forum_user_ids correspond. For P-C and P-Set the nodes are the same but do not rely on odering.
m2.df<-m2.df[rownames(m1.df),]
#eliminate isolates (see markdown text above).
#done before cbind so each can be used separately
m2.df<-m2.df[m1.df[,"sqrt_degree1"]!=0,]
m1.df<-m1.df[m1.df[,"sqrt_degree1"]!=0,]
#
m.df<-cbind(m1.df,m2.df)
plot(m.df[,2:7], main="Correlation of Square Root of Centrality Measures")

# attach(m.df)
# iplot(sqrt_betweenness1, sqrt_betweenness2)
# iplot(sqrt_degree1, sqrt_degree2)#
# iplot(sqrt_betweenness1, sqrt_degree1)
# iplot(sqrt_betweenness2, sqrt_degree2)
# ibar(role)
```

Correlation
---------

Spearman correlation is used; a non-linear but monotonic relationship will appear as a correlation. Spearman rank correlation is appropriate for skew distributions but I have no confidence that the coefficients are quantitatively meaningful. 

```{r results='asis'}
kable(cor(m.df[,-1], use="complete.obs", method="spearman"))
```

Principal Component Analysis
---------
The variance is normalised to 1 for each variable (scale. = T in prcomp). Square root measures are used as before.

NB: PCA is sensitive to outliers so should inspect-and-remove extreme cases (typically "introductions" forums)

Notes on interpretation:
* on the whole, we look at PCA here to summarise non-independence between the centrality measures rather than to select primary components.
* when there are many components, selection often chooses PCs to th eleft of the elbow point in the scree plot. alternatively, PCs with eigenvalues >1 may be selected; these account for more variance than the original variables.

```{r echo=echo.flag}
#prcomp has a bug such that na.action is ignored unless the forumula method is used
pca1 = prcomp(formula=~., data=m1.df[,-1], na.action=na.omit, scale. = TRUE)
plot(pca1, main =paste("Scree Plot for", tie.type.1))
paste("Percentage of variance for PCs in order: ",paste(round(100*(pca1$sdev^2)/sum(pca1$sdev^2),1),collapse="%, "),"%",sep="")
pca1$rotation   # loadings
biplot(pca1, xlabs=m1.df[,1],cex=0.5, main=tie.type.1)
```

```{r echo=echo.flag}
pca2 = prcomp(formula=~., data=m2.df, na.action=na.omit, scale. = TRUE)
plot(pca2)
paste("Percentage of variance for PCs in order: ",paste(round(100*(pca2$sdev^2)/sum(pca2$sdev^2),1),collapse="%, "),"%",sep="")
pca2$rotation   # loadings
biplot(pca2, xlabs=m1.df[,1],cex=0.5, main=tie.type.2)
```

```{r echo=FALSE}
# saves useful data
name<-paste("Tie Type Correlation",courseID,sep=" - ")
fname<-paste("data/",name, ".RData", sep="")
metadata<-list(project=basename(getwd()), origin=name, created=date())
save(list=c("m1.df", "m2.df", "sim1.df", "sim2.df", "tie.type.1", "tie.type.2", "pca1","pca2"),
     file=fname)
cat(paste("Saved to",fname,"\r\n"))
```
